---
title: "PS109_Section 11_raw"
author: "Johanna Reyes O."
date: "2023-04-18"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(stats)
```

## Basics recap, using the "anes_latino" data 

Today, we will be using the "prevote_voted", "trustgov_trustgrev" and the "dem_hisptyp_1st" variable.

*The "prevote_voted" notes whether a responded voted in the last General Election. We will need to subset the data so that we only look into 'yes' and 'no' responses (1, 0).

*The trustgov_trustgrev variable stands for trust in governemnt. We will also need to subset the data so we only use responses within the 1:5 range (always to never), ignoring those who responded -1 (inapplicable), -8 (don't know), or -9 (refused).

### 1. How do we count total number of observations (i.e. the total number of respondents) within a dataframe?

```{r}
#read data

#total observations

```
### 2. How do we know for how many respondents we have information about one variable? (check first prevote_voted, then dem_hisptyp_1st) 

```{r}
#info for origin

#info for polengagement 

```
### 3. How do we know for how many respondents we have information for two variables? (check for dem_hisptype_1st AND prevote_voted)

```{r}
sum(!is.na(y) & !is.na(x)) 

```
### 4. Compute a bar plot with the average value of trust in government (in the y-axis) by first mention of Hispanic origin (in the x-axis). Order the bars from the group with the highest to the lowest trust in government.
* Note: when computing statistics, for example, a mean, we need to include "na.rm=TRUE" to exclude missing cases. Otherwise, we will get NA as the value of the mean.

```{r}

ggplot(data, aes(x, y)) +
    geom_col()

#Check if we have missing data for trustgov_trustgrev

sum(is.na(x))

#now with the mean
ggplot(data, aes(x, mean(y, na.rm = TRUE))) +
    geom_col()

#now we want to subset the data to only look at the values we care about (1-7 for 'dem_hisptyp_1st' and 1-5 for 'trustgov_trustgrev')

data %>%
  filter(x %in% c(1:7), y %in% c(1:5)) %>%
  group_by(x) %>%
  summarize(mean_y = mean(y, na.rm = TRUE)) %>%
  #now apply ggplot
  ggplot(aes(x, mean_y)) +
    geom_col()

#now reorder by decreasing mean trust levels, add fill color, and labels for axes and legend
data %>%
  filter(x %in% c(1:7), y %in% c(1:5)) %>%
  group_by(x) %>%
  summarize(mean_y = mean(y, na.rm = TRUE)) %>%
  #now apply ggplot
  ggplot(aes(reorder(x, y), y, fill = y)) +
    geom_col()+
  
  #rename all labels
   ylab("label")+ #x-axis
    xlab("label") + #y-axis
    labs(fill = "label")+ #label for filling colors
  scale_x_discrete(labels = c("label1", "label2"))+ #labels for dem_hisptyp categories. Fist look at order of the bars on the plot, and then add the labels in that order. 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+ #this angles text
  theme_light()+ # adds a theme
  geom_text(aes(label = (y)), vjust=-0.3, size=3.5) #adds labels to each bar so we know the exact value

#For Labels
#1. Mexican
#2. Mexican-American 
#3. Chicano
#4. Puerto Rican 
#5. Cuban 
#6. Cuban-American
#7. Other
```
*Note: in the assignment, the origin values are not numeric!!!! 

### 5. 

* What is the average value of trust in government for Puerto Ricans?
* What is the average value of trust in government for Chicanx? 

## e. Compare the distribution (aka density) of trust in government between Cubans and Puerto Rican respondents.

* fill = as.factor(X): This maps the X variable to the fill color of the density plot. The as.factor() function is used to convert the X variable to a factor, which is required for fill colors to work properly in ggplot2.

* color = as.factor(X): This maps the X variable to the line color of the density plot. The as.factor() function is used to convert the X variable to a factor, which is required for line colors to work properly in ggplot2.

* alpha = 0.5: This sets the transparency of the density plot to 50%, making it easier to see overlapping areas of the plot.
```{r}
#trustgov_trustgrev and dem_hisptyp_1st

  ggplot(data, aes(x = x, fill = factor(y))) +
  geom_density(alpha = 0.7)
   
#if we need to remove nas: two ways

ggplot(data = na.omit(data), aes(x = x, fill = factor(y))) +
  geom_density(alpha = 0.7)

#or

data %>%
  filter(x %in% c(, )) %>%
  ggplot(aes(x = x, fill = factor(y, labels = c("label1", "label2")))) +
  geom_density(alpha = 0.5)+
   #scale_fill_manual(values = c("#1b9e77", "#d95f02"))+ 
#add appropriate labels


```

### 6. Compute a scatter plot displaying the correlation between Hispanic origin (in the x-axis) and trust in government (y-axis). In your plot, make sure to include a linear regression line summarizing the correlation.
```{r}
ggplot(data, aes(x = x, y = y))+
  geom_point()+
 geom_smooth(method=lm) 
#repeat but now subset to get rid of unwanted values (we only want to look at 1:5 for trust and 1:7 for hispanic type), add appropriate labels for axes and for hispanic type categories

data %>%
  filter(x %in% c(:), y %in% c(:))%>%
ggplot(aes(x = x, y = y))+
  geom_point()+
 geom_smooth(method=lm)

#How can we look at the actual regression estimation?

#summary(lm(y~x, data = data))
```
### 7. Now, compute a scatter plot displaying the correlation between Hispanic origin (in the x-axis) and trust in government, identifying with a different color Hispanics who voted in the general election from those who did not vote. Make sure to include a linear regression line summarizing the correlation for each of these two groups (i.e. yes/no voted).

```{r}
ggplot(data, aes(x = x, y = y, color=z)) +
  geom_point() +
  geom_smooth(method=lm)

#filter for the values we want 

data %>%
  filter(x %in% c(:), y %in% c(:), z %in% c(,)) %>%
  group_by(z) %>%
ggplot(aes(x = x, y = y, color= factor(z, labels = c("label1", "label2")))) +
  geom_point() +
  geom_smooth(method=lm)#Add other pertinent labels

#or, instead of filtering, you can omit na (note that this is not applicable for our anes dataset)

#ggplot(data = na.omit(anes), aes(x = dem_hisptyp_1st, y = trustgov_trustgrev, color=as.factor(prevote_voted))) +
  #geom_point() +
  #geom_smooth(method=lm)
 

```



